---
title: "Untitled"
format: html
---

Script to (geometrically) clean and combine the restoration plot data

```{r}
library(data.table)
library(tidyverse)
library(dplyr)
library(sf)
library(magrittr)
library(ggplot2)
library(lwgeom) 
```

# Select country to do the analysis for
```{r}
country <- 'Senegal'
```


# Import data 
```{r}
cohort.tree.tp <- data.table(read_csv(paste0('data/plot_data/', country, '_tp_species.csv')))
farmer.plot.info.tp <- data.table(read_csv(paste0('data/plot_data/', country, '_tp_profile.csv')))
plots.csv.tp <- data.table(read_csv(paste0('data/plot_data/', country, '_tp_polygon.csv')))
cohort.tree.fmnr <- data.table(read_csv(paste0('data/plot_data/', country, '_fmnr_species.csv')))
farmer.plot.info.fmnr <- data.table(read_csv(paste0('data/plot_data/', country, '_fmnr_profile.csv')))
plots.csv.fmnr <- data.table(read_csv(paste0('data/plot_data/', country, '_fmnr_polygon.csv')))

cohort.tree.tp$type <- 'tp'
farmer.plot.info.tp$type <- 'tp'
plots.csv.tp$type <- 'tp'
cohort.tree.fmnr$type <- 'fmnr'
farmer.plot.info.fmnr$type <- 'fmnr'
plots.csv.fmnr$type <- 'fmnr'

# Make tp and fmrn datasets consistent
farmer.plot.info.fmnr$fmnr_start <- farmer.plot.info.fmnr$fmnr_start_date
farmer.plot.info.tp$fmnr_start <- NA

names(cohort.tree.tp)[names(cohort.tree.tp) == 'tp_latitude'] <- 'latitude'
names(cohort.tree.tp)[names(cohort.tree.tp) == 'tp_longitude'] <- 'longitude'
names(cohort.tree.tp)[names(cohort.tree.tp) == 'tp_altitude'] <- 'altitude'
names(cohort.tree.tp)[names(cohort.tree.tp) == 'tp_accuracy'] <- 'accuracy'
                    
names(cohort.tree.fmnr)[names(cohort.tree.fmnr) == c('tree_latitude', 'tree_longitude',
                                                     'tree_altitude','tree_accuracy')] <- c(
                                                       'latitude','longitude','altitude','accuracy')

# Combine the tp and fmnr datasets
cohort.tree <- rbind(cohort.tree.tp, cohort.tree.fmnr, fill=TRUE)
farmer.plot.info <- rbind(farmer.plot.info.tp, 
                          farmer.plot.info.fmnr[,-c('id', 'land_fenced', 'species_number_start', 'fmnr_start_date')])
plots.csv <- rbind(plots.csv.tp, plots.csv.fmnr[,-c('id')])

rm(cohort.tree.tp, farmer.plot.info.tp, plots.csv.tp,
   cohort.tree.fmnr, farmer.plot.info.fmnr, plots.csv.fmnr)
#### End import
```


```{r}
sf::sf_use_s2(FALSE)
# Re-order plot.info data
farmer.plot.info <- farmer.plot.info[order(farmer.plot.info$farmerID),]

#### MAKE PLOT DATA SPATIAL AND ASSIGN ID
# Make sf from plot.geom
plots.attr <- plots.csv[,-c("area_polygon")] # select all columns except for 'Plot"
geometry <- st_as_sfc(plots.csv$area_polygon, crs = 4326)
plots.spat <- st_sf(plots.attr, geometry)

# Clean plot.spat file
names(plots.spat)[names(plots.spat) == 'plotID'] <- 'Plot_id'
plots.spat <- plots.spat[plots.spat$area_msq!=0,] # Remove where area is 0

# Assign plotID to datasets
plotID <- seq(1, nrow(plots.spat))
plots.spat$plotID <- seq(1, nrow(plots.spat))

# Convert multipolygon into single polygons
plots.spat <- st_cast(plots.spat, "POLYGON")
#Calculate size of (potential) new polygons
plots.spat$area <- as.numeric(st_area(plots.spat)/10000) # in ha
plots.spat <- plots.spat[plots.spat$area>0,] # remove plots with area = 0
# Order on polygon size and then remove duplicates (smallest)
plots.spat <- plots.spat[order(plots.spat$plotID),]
#plots.spat <- plots.spat[-c(which(duplicated(plots.spat$plotID, fromLast=F))),]
# Make geometries valid
plots.spat <- st_make_valid(plots.spat)
```


# Exclude and merge overlapping geometries
```{r}
# Set buffer of 0m to remove intersection issues
sf::sf_use_s2(FALSE)
plots.spat <- st_buffer(plots.spat, 0)
# Create empty df to fill
for (type in unique(plots.spat$type)){
  print(type)
  plots_type <- plots.spat[plots.spat$type==type,]
  
  for (pl in unique(plots_type$plotID)){
    
    intersect_rows <- st_intersection(plots_type[plots_type$plotID==pl,], plots_type)
    print(pl)
    # Check whether plot has not yet been excluded yet
    if (nrow(intersect_rows)!=0){
      
      intersect_rows$OverlapArea <- st_area(intersect_rows)
      other_plots <- intersect_rows[intersect_rows$plotID.1!=pl,]
      sum_overlap <- sum(other_plots$OverlapArea)
      area_plot_interest <- intersect_rows$OverlapArea[intersect_rows$plotID.1==pl]
      perc_overlap <- (sum_overlap/area_plot_interest)*100
      # Set threshold for when percentage overlap is too high (now: 60%)
      if (as.integer(perc_overlap)>as.integer(60)){
        # Find plot with largest overlap
        largest_overlap_plotID <- other_plots$plotID.1[other_plots$OverlapArea==max(other_plots$OverlapArea)]
        # Select row of largest overlapping plot
        largest_overlap_row <- plots_type[plots_type$plotID==largest_overlap_plotID,]
        # Join the plot of interest and the largest overlapping plot
        join_interest_largest <- rbind(plots_type[plots_type$plotID==pl,], largest_overlap_row)
        # Select the biggest of the two
        largest_plot_row <- join_interest_largest[join_interest_largest$area_msq==
                                                    max(join_interest_largest$area_msq),]
        # Join the geometries with the largest-overlap plot
        joined_geom <- st_union(plots_type[plots_type$plotID==pl,], largest_plot_row, is_coverage=T)
        plots_type$geometry[plots_type$plotID==largest_plot_row$plotID] <- joined_geom$geometry
        # Exclude overlapping plot from df
        smallest_plot_ID <- join_interest_largest$plotID[join_interest_largest$area_msq==
                                                           min(join_interest_largest$area_msq)]
        plots.spat <- plots.spat[plots.spat$plotID!=smallest_plot_ID,]
        
        print(paste(type, 'polygon excluded'))
      }
    }
    else{print('Plot already exlcuded')}
  }
}
st_write(obj=plots.spat, dsn=paste0('output/plot_data/', country, '/', country, '_plots_filtered.GeoJSON'), driver='GeoJSON')
plots.spat <- st_read(dsn=paste0('output/plot_data/', country, '/', country, '_plots_filtered.GeoJSON'))
plots.backup <- plots.spat
```


Make plots.all similar to plots.spat so line 98-147 do have to be ran (takes a while). Note: this is a fix, not origional code.
```{r}
# Import Ghana filtered to get the column names that are normally at this point
plots.spat <- st_read(dsn=paste0('output/plot_data/', 'Ghana', '/', 'Ghana', '_plots_filtered.GeoJSON'))

# Only select the columns in plots that are present in plots.spat
plots <- st_read(paste0('output/plot_data/', country, '/', country, '_plots_all.GeoJSON'))
plots$country <- country
names(plots)[names(plots) == 'Hectare'] <- 'area_polygon'

# Check if all columns correspond
names(plots.spat)[-2] %in% names(plots)

plots.spat <- plots[,names(plots.spat)[-2]]
```


# Indicate if multiple plots per farmer ID so plots can be assigned a unique plot ID
```{r}
# for plots.spat dt
dupl.index <- which(duplicated(plots.spat$farmerID) | duplicated(plots.spat$farmerID, fromLast=TRUE)==T)
dupl.unique <- unique(plots.spat$farmerID[dupl.index])

plots.spat$MultiplePlots <- 0
for (i in dupl.unique){
  num.plots.farm <- nrow(plots.spat[plots.spat$farmerID==i,])
  index.plots.farm <- which(plots.spat$farmerID==i)
  for (pl in 1:length(index.plots.farm)){
    plots.spat$MultiplePlots[index.plots.farm[pl]] <- pl
  }
}
plots.spat$FarmerPlotID <- NA
plots.spat$FarmerPlotID <- paste0(plots.spat$farmerID, '_', plots.spat$MultiplePlots)

# for farmer.plot.info dt
dupl.index <- which(duplicated(farmer.plot.info$farmerID) | duplicated(farmer.plot.info$farmerID, fromLast=TRUE)==T)
dupl.unique <- unique(farmer.plot.info$farmerID[dupl.index])

farmer.plot.info$MultiplePlots <- 0
for (i in dupl.unique){
  num.plots.farm <- nrow(farmer.plot.info[farmer.plot.info$farmerID==i,])
  index.plots.farm <- which(farmer.plot.info$farmerID==i)
  for (pl in 1:length(index.plots.farm)){
    farmer.plot.info$MultiplePlots[index.plots.farm[pl]] <- pl
  }
}
farmer.plot.info$FarmerPlotID <- NA
farmer.plot.info$FarmerPlotID <- paste0(farmer.plot.info$farmerID, '_', farmer.plot.info$MultiplePlots)
```


# Combine cohort and plot data based on farmerID or spatial overlap
```{r}
# Remove the rows with NAs in the longitude, latitude or farmerID
cohort.tree <- na.omit(cohort.tree, cols=c('longitude', 'latitude','farmerID'))
# Make the cohort.tree observations spatial
cohort.spat <- st_as_sf(cohort.tree, coords = c("longitude", "latitude"), crs = 4326)

# Assign plotID to tree data based on location tree
sf::sf_use_s2(FALSE) # turn of s2 to avoid error in st_intersection

# 1 lat = ~ 111111 m at the equator --> 5m = 0.000045 (7m = 0.000063)
plots.buf <- st_buffer(plots.spat, 0.000063)

cohort.spat$plotID <- NA
for (pl in 1:nrow(plots.buf)){
  print(pl)
  # If on plot per farmer --> merge on farmerID
  if(plots.buf$MultiplePlots[pl]==0){
    match_farmerID <- cohort.spat$farmerID[cohort.spat$farmerID==plots.buf$farmerID[pl]]
    cohort.spat$plotID[cohort.spat$farmerID==match_farmerID] <- plots.buf$plotID[pl]
  }
  # If multiple plots per farmer --> merge based on spatial location
  else{
    trees.in.plot.index <- which(st_within(cohort.spat, plots.buf[pl,], sparse=FALSE)==TRUE)
    trees.in.plot <- cohort.spat[c(trees.in.plot.index),]
    
    if (nrow(trees.in.plot) >= 1) {
      for (t in 1:nrow(trees.in.plot)){
        if (trees.in.plot$farmerID[t]==plots.buf$farmerID[pl]){
          cohort.spat$plotID[trees.in.plot.index[t]] <- plots.buf$plotID[pl]
        }
      }
    }
  }
}
```


# Assign date to cohort and farmer data
```{r}
# TO COHORT DATA
months <- c('Jan'=1, 'Feb'=2, 'Mar'=3, 'Apr'=4,'May'=5,'Jun'=6,'Jul'=7,'Aug'=8,'Sep'=9,
            'Oct'=10,'Nov'=11,'Dec'=12)
cohort.spat$plant_date <- as.Date(ISOdate(2000,1,1))

for (i in 1:nrow(cohort.spat)){
  print(i)
  
  date_string <- cohort.spat$date_planted[i]
  split_list <- strsplit(date_string, '/')
  
  cohort.spat$plant_date[i] <- as.Date(ISOdate(as.integer(split_list[[1]][2]), 
                                               #as.integer(months[split_list[[1]][1]]),
                                               as.integer(split_list[[1]][1]), 
                                               1))
}
# TO FARMER DATA
farmer.plot.info$plant_date_fmnr <- as.Date(ISOdate(2000,1,1))
for (i in 1:nrow(farmer.plot.info)){
  print(i)
  
  date_string <- farmer.plot.info$fmnr_start[i]
  split_list <- strsplit(date_string, '/')
  
  farmer.plot.info$plant_date_fmnr[i] <- as.Date(ISOdate(as.integer(split_list[[1]][2]), 
                                               as.integer(split_list[[1]][1]), 
                                               1))
}
```


# Summarise cohort information at plot level and indicate first cohort (for merging purposes)
```{r}
cohort.info <- data.table(st_drop_geometry(cohort.spat)) # remove geometry

####### NEW METHOD --> Check if management practise has been applied at the plot level in any cohort
cohort.info <- cohort.info %>% 
    mutate(mgt_prunning = if_else(mgt_prunning == 'yes', 1, 0),
        mgt_fencing = if_else(mgt_fencing == 'yes', 1, 0),
        mgt_weeding = if_else(mgt_weeding == 'yes', 1, 0),
        mgt_watering = if_else(mgt_watering == 'yes', 1, 0),
        mgt_organic_fertilizer = if_else(mgt_organic_fertilizer == 'yes', 1, 0),
        mgt_thinning = if_else(mgt_thinning == 'yes', 1, 0),
        mgt_pollarding_lopping = if_else(mgt_pollarding_lopping == 'yes', 1, 0),
        mgt_coppicing = if_else(mgt_coppicing == 'yes', 1, 0),
        mgt_others = if_else(mgt_others == 'yes', 1, 0))

cohort.info <- cohort.info %>% 
  group_by(plotID) %>% 
  mutate(mgt_prunning = if_else(sum(mgt_prunning, na.rm=TRUE) > 0, 1, 0),
            mgt_fencing = if_else(sum(mgt_fencing, na.rm=TRUE) > 0, 1, 0),
            mgt_weeding = if_else(sum(mgt_weeding, na.rm=TRUE) > 0, 1, 0),
            mgt_watering = if_else(sum(mgt_watering, na.rm=TRUE) > 0, 1, 0),
            mgt_organic_fertilizer = if_else(sum(mgt_organic_fertilizer, na.rm=TRUE) > 0, 1, 0),
            mgt_thinning = if_else(sum(mgt_thinning, na.rm=TRUE) > 0, 1, 0),
            mgt_pollarding_lopping = if_else(sum(mgt_pollarding_lopping, na.rm=TRUE) > 0, 1, 0),
            mgt_coppicing = if_else(sum(mgt_coppicing, na.rm=TRUE) > 0, 1, 0),
            mgt_others = if_else(sum(mgt_others, na.rm=TRUE) > 0, 1, 0)) %>% 
  data.table()

# Indicate the first cohort at each plot (for merging purposes)
cohort.info$FirstCohortPlotID <- 0
# Select plotID only for the first (based on date) cohort in plot
for (i in unique(cohort.info$plotID)){
  print(i)
  cohort_rows <- cohort.info[cohort.info$plotID==i,]
  # Check if there is a plant date available
  if(length(which(!is.na(cohort_rows$plant_date))) == nrow(cohort_rows)){
      first_plant <- which(cohort_rows$plant_date==min(cohort_rows$plant_date))[1]
      cohort.info[cohort.info$plotID==i,]$FirstCohortPlotID[first_plant] <- i
  }
  # If no plant date available, fill FirstCohortPlotID of first row
  else{
    cohort.info[cohort.info$plotID==i,]$FirstCohortPlotID[1] <- i
  }
}
```


# Merge datasets on a plot level
```{r}
# Add farmer info
farmer.plot.info.geom <- merge(farmer.plot.info[,c('country', 'county_region', 'district_commune_woreda', 'own_land', 'community_land', 'government_land', 'mosque_church_land', 'school_land', 'other_land', 'crops', 'croplist', 'unit', 'landsize', 'FarmerPlotID', 'plant_date_fmnr')], plots.spat[,c('farmerID', 'Plot_id', 'area_msq', 'plotID', 'MultiplePlots', 'FarmerPlotID','type')], by='FarmerPlotID', all.x=FALSE, all.y=TRUE)
farmer.plot.info.spat <- st_as_sf(farmer.plot.info.geom) # Make spatial

# Add cohort info
plots.all <- merge(farmer.plot.info.spat, cohort.info[,c('species_name', 'plant_date', 'number_planted', 'number_survived', 'woodlot', 'internal_boundary', 'external_boundary', 'garden', 'crop_field', 'pasture_grassland','fallow_bushland', 'other_sites', 'mgt_prunning', 'mgt_fencing', 'mgt_weeding', 'mgt_watering','mgt_organic_fertilizer', 'FirstCohortPlotID','use_firewood', 'local_name','mgt_thinning', 'mgt_pollarding_lopping','mgt_coppicing','mgt_others')], by.x='plotID', by.y = 'FirstCohortPlotID', all.x=TRUE, all.y=FALSE)
# If fmnr cohort information is available add
#plots.all$mgt_others <- NA
#c('local_name','mgt_thinning','mgt_pollarding_lopping','mgt_coppicing','mgt_others')

# Merge TP and FMNR plant date together
plots.all$plant_date <- fifelse(is.na(plots.all$plant_date), plots.all$plant_date_fmnr, plots.all$plant_date)

# Make sure country variable is correct
plots.all$country <- country
```


Check if management information is present at fmnr plots (new code)
```{r}
# Check is length plots.all is similar to length plots
nrow(plots.all) == nrow(plots)

plots.all %>% 
    st_drop_geometry() %>% 
    filter(type == 'fmnr') %>% 
    group_by(mgt_prunning) %>% 
    summarise(n = n())


plots_final %>% 
    st_drop_geometry() %>% 
    filter(type == 'tp') %>% 
    group_by(mgt_prunning) %>% 
    summarise(n = n())

plots %>% 
    #filter(type == 'fmnr') %>% 
    group_by(mgt_prunning) %>% 
    summarise(n = n())

country <- 'Kenya'
# Check difference in length between plots.all and plots.centroid
plots.all <- read_csv(paste0('output/plot_data/', country, '/', country, '_plots_all.csv'))
plots.all <- st_read(paste0('output/plot_data/', country, '/', country, '_plots_all.GeoJSON'))
plot.centroid <- read_csv(paste0('output/plot_data/', country, '/', country, '_plot_centroid.csv'))

nrow(plots.all) == nrow(plot.centroid)
```


# Clean combined df
```{r}
#### CLEAN COMBINED DF
sf::sf_use_s2(FALSE)
plots.all <- st_buffer(plots.all, 0)
# Convert multipolygon into single polygons
plots.all <- st_cast(plots.all, "POLYGON")
plots.all <- st_make_valid(plots.all)
# Recalculate the area after the spatial merges
plots.all$Hectare <- as.numeric(st_area(plots.all)/10000) # in ha
plots.all <- plots.all[plots.all$Hectare>0,] # remove plots with area = 0
# Order on polygon size and then remove duplicates (smallest)
plots.all <- plots.all[order(plots.all$Hectare),]
#plots.all <- plots.all[-c(which(duplicated(plots.all$plotID, fromLast=FALSE))),]
```


# Change the geometries of the plot polygons
```{r}
# Centroid
plot.centroid <- st_centroid(plots.all, of_largest_polygon=FALSE)
# Multipolygon
centroid_multi <- plot.centroid %>% 
  st_combine()
```


# Write the files as geojson and csv
```{r}
st_write(obj=plots.all, dsn=paste0('output/plot_data/', country, '/', country, '_plots_all.GeoJSON'), driver='GeoJSON')
st_write(obj=plots.all, dsn=paste0('output/plot_data/', country, '/', country, '_plots_all.csv'), driver='CSV')
st_write(obj=plot.centroid, dsn=paste0('output/plot_data/', country, '/', country, '_plot_centroid.GeoJSON'), driver='GeoJSON')
st_write(obj=plot.centroid, dsn=paste0('output/plot_data/', country, '/', country, '_plot_centroid.csv'), driver='CSV')
st_write(obj=centroid_multi, dsn=paste0('output/plot_data/', country, '/', country, '_centroid_multi.GeoJSON'), driver='GeoJSON')
st_write(obj=centroid_multi, dsn=paste0('output/plot_data/', country, '/', country, '_centroid_multi.csv'), driver='CSV')
```
